---
title: "Exploring data I"
subtitle: "Lecture 4"
date: "2024-09-10"
format: 
  live-revealjs: 
    output-file: 04-exploring-data-I-slides.html
    webr:
      cell-options:
        autorun: false
webr:
  packages:
    - tidyverse
    - ggthemes
    - usdata
---

<!-- begin: webr fodder -->

{{< include ../_extensions/r-wasm/live/_knitr.qmd >}}

```{webr}
#| edit: false
#| echo: false
#| output: false
options(width = 60)
```

<!-- end: webr fodder -->

<!-- begin: ae definition -->

```{r}
#| include: false
todays_ae <- "ae-04-gerrymander-explore-I"
```

<!-- end: ae definition -->

# Warm-up

## While you wait... {.smaller}

Prepare for today's application exercise: **`{r} todays_ae`**

::: appex
-   Go to your `ae` project in RStudio.

-   Make sure all of your changes up to this point are committed and pushed, i.e., there's nothing left in your Git pane.

-   Click Pull to get today's application exercise file: *`{r} paste0(todays_ae, ".qmd")`*.

-   Wait till the you're prompted to work on the application exercise during class before editing the file.
:::

## Announcements {.smaller}

::: incremental
-   **Weekend office hours** in Old Chem – bring your Duke Card, you'll need to swipe/tap in to get in

-   **AEs are due by the end of class** (precisely, by 2 pm) – done = at least one commit + push

-   **Coding principles bonus office hour**, RSVP at <https://forms.gle/mufcsHnPXejZbkdT8> – Thursday, 9/12 at 7:30 pm at Old Chem 116

-   **RStudio visual editor acting up?**

    -   Using Chrome – make sure to update!

    -   Using another browser or update doesn't solve the problem, come by my office hours to diagnose

-   New resource: **Code along videos**

-   Any **questions about lab** before we get started?
:::

# Exploratory data analysis

## Packages {.smaller}

-   For the data: [**usdata**](https://openintrostat.github.io/usdata/)

```{r}
library(usdata)
```

-   For the analysis: [**tidyverse**](https://www.tidyverse.org/packages/) and [**ggthemes**](https://jrnold.github.io/ggthemes/)

```{r}
library(tidyverse)
library(ggthemes)
```

## Data: `gerrymander` {.smaller}

```{r}
gerrymander
```

## What is gerrymandering?

<https://www.washingtonpost.com/business/wonkblog/gerrymandering-explained/2016/04/21/e447f5c2-07fe-11e6-bfed-ef65dff5970d_video.html>

## Data: `gerrymander` {.scrollable .smaller}

::: question
What is a good first function to use to get to know a dataset?
:::

```{webr}

```

## Data: `gerrymander` {.smaller}

-   Rows: Congressional districts

-   Columns:

    -   Congressional district and state

    -   2016 election: winning party, % for Clinton, % for Trump, whether a Democrat won the House election, name of election winner

    -   2018 election: winning party, whether a Democrat won the 2018 House election

    -   Whether a Democrat flipped the seat in the 2018 election

    -   Prevalence of gerrymandering: low, mid, and high

## Variable types

::: columns
::: column
::: xsmall
| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | [categorical, ID]{.fragment}       |
| `last_name`  | [categorical, ID]{.fragment}       |
| `first_name` | [categorical, ID]{.fragment}       |
| `party16`    | [categorical]{.fragment}           |
| `clinton16`  | [numerical, continuous]{.fragment} |
| `trump16`    | [numerical, continuous]{.fragment} |
| `dem16`      | [categorical]{.fragment}           |
| `state`      | [categorical]{.fragment}           |
| `party18`    | [categorical]{.fragment}           |
| `dem18`      | [categorical]{.fragment}           |
| `flip18`     | [categorical]{.fragment}           |
| `gerry`      | [categorical, ordinal]{.fragment}  |
:::
:::

::: column
```{webr}
gerrymander |>
  select(___)
```
:::
:::

# Univariate analysis

## Univariate analysis

Analyzing a single variable:

::: incremental
-   Numerical: histogram, box plot, density plot, etc.

-   Categorical: bar plot, pie chart, etc.
:::

## Histogram - Step 1 {.smaller .scrollable}

```{r}
ggplot(gerrymander)
```

## Histogram - Step 2 {.smaller .scrollable}

```{r}
ggplot(gerrymander, aes(x = trump16))
```

## Histogram - Step 3 {.smaller .scrollable}

```{r}
#| code-line-numbers: "2"
ggplot(gerrymander, aes(x = trump16)) +
  geom_histogram()
```

## Histogram - Step 4 {.smaller .scrollable}

```{webr}
ggplot(gerrymander, aes(x = trump16)) +
  geom_histogram(binwidth = ___)
```

## Histogram - Step 5 {.smaller .scrollable}

```{r}
#| code-line-numbers: "3-8"
ggplot(gerrymander, aes(x = trump16)) +
  geom_histogram(binwidth = 5) +
  labs(
    title = "Percent of vote received by Trump in 2016 Presidential Election",
    subtitle = "From each Congressional District",
    x = "Percent of vote",
    y = "Count"
  )
```

## Box plot - Step 1 {.smaller .scrollable}

```{r}
ggplot(gerrymander)
```

## Box plot - Step 2 {.smaller .scrollable}

```{r}
ggplot(gerrymander, aes(x = trump16))
```

## Box plot - Step 3 {.smaller .scrollable}

```{r}
#| code-line-numbers: "2"
ggplot(gerrymander, aes(x = trump16)) +
  geom_boxplot()
```

## Box plot - Alternative Step 2 + 3 {.smaller .scrollable}

```{webr}
ggplot(gerrymander, aes(___ = trump16)) +
  geom_boxplot()
```

## Box plot - Step 4 {.smaller .scrollable}

```{r}
#| code-line-numbers: "3-8"
ggplot(gerrymander, aes(x = trump16)) +
  geom_boxplot() +
  labs(
    title = "Percent of vote received by Trump in 2016 Presidential Election",
    subtitle = "From each Congressional District",
    x = "Percent of vote",
    y = NULL
  )
```

## Density plot - Step 1 {.smaller .scrollable}

```{r}
ggplot(gerrymander)
```

## Density plot - Step 2 {.smaller .scrollable}

```{r}
ggplot(gerrymander, aes(x = trump16))
```

## Density plot - Step 3 {.smaller .scrollable}

```{r}
#| code-line-numbers: "2"
ggplot(gerrymander, aes(x = trump16)) +
  geom_density()
```

## Density plot - Step 4 {.smaller .scrollable}

```{webr}
ggplot(gerrymander, aes(x = trump16)) +
  geom_density(color = ___)
```

## Density plot - Step 5 {.smaller .scrollable}

```{webr}
ggplot(gerrymander, aes(x = trump16)) +
  geom_density(color = "firebrick", fill = ___)
```

## Density plot - Step 6 {.smaller .scrollable}

```{webr}
ggplot(gerrymander, aes(x = trump16)) +
  geom_density(color = "firebrick", fill = "firebrick1", alpha = ___)
```

## Density plot - Step 7 {.smaller .scrollable}

```{webr}
ggplot(gerrymander, aes(x = trump16)) +
  geom_density(color = "firebrick", fill = "firebrick1", alpha = 0.5, linewidth = ___)
```

## Density plot - Step 8 {.smaller .scrollable}

```{r}
#| code-line-numbers: "3-8"
ggplot(gerrymander, aes(x = trump16)) +
  geom_density(color = "firebrick", fill = "firebrick1", alpha = 0.5, linewidth = 2) +
  labs(
    title = "Percent of vote received by Trump in 2016 Presidential Election",
    subtitle = "From each Congressional District",
    x = "Percent of vote",
    y = "Density"
  )
```

## Summary statistics

```{webr}
#| autorun: true
gerrymander |>
  summarize(
    mean = mean(trump16),
    median = median(trump16)
  )
```

## Distribution of votes for Trump in the 2016 election {.smaller}

::: question
Describe the distribution of percent of vote received by Trump in 2016 Presidential Election from Congressional Districts.
:::

-   Shape: [The distribution of votes for Trump in the 2016 election from Congressional Districts is **unimodal and left-skewed**.]{.fragment}

-   Center: [The percent of vote received by Trump in the 2016 Presidential Election from a **typical** Congressional Districts is 48.7%.]{.fragment}

-   Spread: [In the **middle 50%** of Congressional Districts, 34.8% to 58.1% of voters voted for Trump in the 2016 Presidential Election.]{.fragment}

-   Unusual observations: [-]{.fragment}

# Bivariate analysis

## Bivariate analysis

Analyzing the relationship between two variables:

::: incremental
-   Numerical + numerical: scatterplot

-   Numerical + categorical: side-by-side box plots, violin plots, etc.

-   Categorical + categorical: stacked bar plots

-   Using an aesthetic (e.g., fill, color, shape, etc.) or facets to represent the second variable in any plot
:::

## Side-by-side box plots

```{r}
#| code-line-numbers: "|2|3-6|4|5|8"
#| output-location: column
ggplot(
  gerrymander, 
  aes(
    x = trump16, 
    y = gerry
    )
  ) +
  geom_boxplot()
```

## Summary statistics

```{webr}
gerrymander |>
  # do the following for each level of gerry
  summarize(
    min = min(trump16),
    q25 = quantile(trump16, 0.25),
    median = median(trump16),
    q75 = quantile(trump16, 0.75),
    max = max(trump16),
  )
```

## Density plots

```{r}
#| code-line-numbers: "|4|5|8"
#| output-location: column
ggplot(
  gerrymander, 
  aes(
    x = trump16, 
    color = gerry
    )
  ) +
  geom_density()
```

## Filled density plots

```{r}
#| code-line-numbers: "6"
#| output-location: column
ggplot(
  gerrymander, 
  aes(
    x = trump16, 
    color = gerry,
    fill = gerry
    )
  ) +
  geom_density()
```

## Better filled density plots

```{r}
#| code-line-numbers: "5"
ggplot(
  gerrymander, 
  aes(x = trump16, color = gerry, fill = gerry)
  ) +
  geom_density(alpha = 0.5)
```

## Better colors

```{r}
#| code-line-numbers: "7-8"
#| output-location: column
ggplot(
  gerrymander, 
  aes(x = trump16, color = gerry, fill = gerry)
  ) +
  geom_density(alpha = 0.5) +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

## Violin plots {.scrollable}

```{r}
#| code-line-numbers: "5"
ggplot(
  gerrymander, 
  aes(x = trump16, y = gerry, color = gerry)
  ) +
  geom_violin() +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

## Multiple geoms {.scrollable}

```{r}
#| code-line-numbers: "6"
ggplot(
  gerrymander, 
  aes(x = trump16, y = gerry, color = gerry)
  ) +
  geom_violin() +
  geom_point() +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

## Multiple geoms {.scrollable}

```{r}
#| code-line-numbers: "6"
ggplot(
  gerrymander, 
  aes(x = trump16, y = gerry, color = gerry)
  ) +
  geom_violin() +
  geom_jitter() +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

## Remove legend {.scrollable}

```{r}
#| code-line-numbers: "9"
ggplot(
  gerrymander, 
  aes(x = trump16, y = gerry, color = gerry)
  ) +
  geom_violin() +
  geom_jitter() +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  theme(legend.position = "none")
```

# Multivariate analysis

## Multivariate analysis {.smaller}

Analyzing the relationship between multiple variables:

::: incremental
-   In general, one variable is identified as the **outcome** of interest

-   The remaining variables are **predictors** or **explanatory variables**

-   Plots for exploring multivariate relationships are the same as those for bivariate relationships, but **conditional** on one or more variables

    -   Conditioning can be done via faceting or aesthetic mappings (e.g., scatterplot of `y` vs. `x1`, colored by `x2`, faceted by `x3`)

-   Summary statistics for exploring multivariate relationships are the same as those for bivariate relationships, but **conditional** on one or more variables

    -   Conditioning can be done via grouping (e.g., correlation between `y` and `x1`, grouped by levels of `x2` and `x3`)
:::

# Application exercise

## `{r} todays_ae` {.smaller}

::: appex
-   Go to your ae project in RStudio.

-   If you haven't yet done so, make sure all of your changes up to this point are committed and pushed, i.e., there's nothing left in your Git pane.

-   If you haven't yet done so, click Pull to get today's application exercise file: *`{r} paste0(todays_ae, ".qmd")`*.

-   Work through the application exercise in class, and render, commit, and push your edits by the end of class.
:::
