---
title: "Linear regression with a multiple predictors II"
subtitle: "Lecture 17"
date: "2024-10-31"
format: 
  live-revealjs: 
    output-file: 17-linear-model-multiple-predictors-II-slides.html
webr:
  cell-options:
    autorun: false
---

<!-- begin: webr fodder -->

{{< include ../_extensions/r-wasm/live/_knitr.qmd >}}

```{webr}
#| edit: false
#| echo: false
#| output: false
options(width = 60)
```

```{r}
#| include: false
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
```

<!-- end: webr fodder -->

<!-- begin: ae definition -->

```{r}
#| include: false
todays_ae <- "ae-14-modeling-penguins-multi"
```

<!-- end: ae definition -->

# Warm-up

## While you wait... {.smaller}

::: appex
-   Go to your `ae` project in RStudio.

-   Make sure all of your changes up to this point are committed and pushed, i.e., there's nothing left in your Git pane.

-   Click Pull to get today's application exercise file: *`{r} paste0(todays_ae, ".qmd")`*.

-   Wait till the you're prompted to work on the application exercise during class before editing the file.
:::

## Announcements

-   Peer evaluations (via TEAMMATES) are due by Friday (tomorrow), 5 pm.
-   

# Application exercise

## Finish up Tuesday's AE

-   Go to your ae project in RStudio.

-   Get back to working on `ae-14-modeling-penguins-multi`

## `{r} todays_ae` {.smaller}

::: appex
-   Go to your ae project in RStudio.

-   If you haven't yet done so, make sure all of your changes up to this point are committed and pushed, i.e., there's nothing left in your Git pane.

-   If you haven't yet done so, click Pull to get today's application exercise file: *`{r} paste0(todays_ae, ".qmd")`*.

-   Work through the application exercise in class, and render, commit, and push your edits.
:::