---
title: "Model selection and overfitting"
subtitle: "Lecture 18"
date: "2024-11-05"
format: 
  live-revealjs: 
    output-file: 18-model-selection-overfitting-slides.html
webr:
  cell-options:
    autorun: false
---

<!-- begin: webr fodder -->

{{< include ../_extensions/r-wasm/live/_knitr.qmd >}}

```{webr}
#| edit: false
#| echo: false
#| output: false
options(width = 60)
```

```{r}
#| include: false
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
```

<!-- end: webr fodder -->

<!-- begin: ae definition -->

```{r}
#| include: false
todays_ae <- "ae-15-modeling-loans"
```

<!-- end: ae definition -->

# Warm-up

## While you wait... {.smaller}

::: appex
-   Go to your `ae` project in RStudio.

-   Make sure all of your changes up to this point are committed and pushed, i.e., there's nothing left in your Git pane.

-   If you missed class last Thursday, pull to get today's application exercise file: *`{r} paste0(todays_ae, ".qmd")`*.

-   Make sure you've completed the "Get to know the data" section of your AE.
:::

## Announcements {.smaller}

-   My office hours this week:
    -   I'll hold them at a modified time: 1-2 pm on Wednesday at Old Chem 213 (in place of Dav's office hours)
    -   Dav will fill in for me 2-4 pm on Wednesday at Old Chem 203B
-   Make sure you're caught up with prepare materials before Thursday's class

# Reminders

## What is the difference between $R^2$ and adjusted $R^2$? {.smaller}

-   $R^2$:

    -   Proportion of variability in the outcome explained by the model.

    -   Useful for quantifying the fit of a given model.

-   Adjusted $R^2$:

    -   Proportion of variability in the outcome explained by the model, with a penalty added for the number of predictors in the model.

    -   Useful for comparing models.

# Application exercise

## Finish up Thursday's AE {.smaller}

::: appex
-   Go to your ae project in RStudio.

-   Get back to working on `ae-15-modeling-loans`
:::

**Goals:**

-   Review prediction and interpretation of model results

-   Review main and interaction effects models

-   Discuss model selection further

## Recap

-   What is the practical difference between a model with parallel and non-parallel lines?

-   What is the definition of R-squared?

-   Why do we choose models based on adjusted R-squared and not R-squared?
