---
title: "AE 08: Durham climate + factors"
---

## Getting started

### Packages

We will use the **tidyverse** package in this application exercise.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
```

### Data

The data come from <https://www.usclimatedata.com/climate/durham/north-carolina/united-states/usnc0192> and are stored as `durham-climate.csv` in the `data` folder.

```{r}
#| label: load-data
#| message: false
durham_climate <- read_csv("data/durham-climate.csv")
```

And let's take a look at the data.

```{r}
#| label: view-data
durham_climate
```

## Reorder

What's wrong with the following plot?

Add your response here.

```{r}
#| label: avg-high-broken
#| fig-width: 8
#| fig-asp: 0.4
#| out-width: 100%
ggplot(
  durham_climate, 
  aes(x = month, y = avg_high_f, group = 1)
  ) +
  geom_line() +
  geom_point(
    shape = "circle filled", size = 2,
    color = "black", fill = "white", stroke = 1
  ) +
  labs(
    x = "Month",
    y = "Average high temperature (F)",
    title = "Durham climate"
  )
```

Fix the plot.

```{r}
#| label: avg-high-fixed
#| fig-width: 8
#| fig-asp: 0.4
#| out-width: 100%
# add code here
```

## Recode and reorder

Update the plot above, coloring points based on season.
Additionally:

-   Make sure the legend is on top of the plot and the seasons appear in the legend in the order they appear in the plot (Winter, Spring, Summer, Fall).

-   Use `"circle filled"` as the `shape` for points, set their `size` to 3 points, outline (`stroke`) to 1 point, and fill them in with the following colors:

    -   Winter: `lightskyblue1`
    -   Spring: `chartreuse3`
    -   Summer: `gold2`
    -   Fall: `lightsalmon4`

```{r}
#| label: avg-high-seasons
#| fig-width: 8
#| fig-asp: 0.4
#| out-width: 100%
# add code here
```

## Pivot

Overlay lines for both high and low temperatures on the same plot.
Additionally, use different colors for the two lines – a darker color for high and a lighter color for low temperatures.

```{r}
#| label: avg-seasons
#| fig-width: 8
#| fig-asp: 0.4
#| out-width: 100%
# add code here
```
