---
title: "AE 02: Bechdel + data visualization"
---

In this mini-analysis, we use the data from the FiveThirtyEight story "The Dollar-And-Cents Case Against Hollywood's Exclusion of Women."

This analysis is about the [Bechdel test](https://en.wikipedia.org/wiki/Bechdel_test), a measure of the representation of women in fiction.

## Getting started

### Packages

We'll use the **tidyverse** package for this analysis.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
```

### Data

The data are stored as a CSV (comma-separated values) file in your repository's `data` folder.
Let's read it from there and save it as an object called `bechdel`.

```{r}
#| label: read-data
#| message: false
bechdel <- read_csv("data/bechdel.csv")
```

### Get to know the data

We can use the `glimpse()` function to get an overview (or "glimpse") of the data.

```{r}
#| label: glimpse
# add code here
```

-   What does each observation (row) in the data set represent?

Each observation represents a \_\_\_.

-   How many observations (rows) are in the data set?

There are \_\_\_ movies in the dataset.

-   How many variables (columns) are in the data set?

There are \_\_\_ columns in the dataset.

### Variables of interest

The variables we'll focus on are the following:

-   `roi`: Return on investment, calculated as the ratio of the gross to budget.
-   `clean_test`: Bechdel test result:
    -   `ok` = passes test
    -   `dubious`
    -   `men` = women only talk about men
    -   `notalk` = women don't talk to each other
    -   `nowomen` = fewer than two women
-   `binary`: Bechdel Test PASS vs FAIL binary

We will also use the `year` of release in data prep and `title` of movie to take a deeper look at some outliers.

There are a few other variables in the dataset, but we won't be using them in this analysis.

## Bechdel test results

### Visualizing data with **ggplot2**

**ggplot2** is the package and `ggplot()` is the function in this package that is used to create a plot.

-   `ggplot()` creates the initial base coordinate system, and we will add layers to that base. We first specify the data set we will use with `data = bechdel`.

```{r}
#| label: plot-base
ggplot(data = bechdel)
```

-   The `mapping` argument is paired with an aesthetic (`aes()`), which tells us how the variables in our data set should be mapped to the visual properties of the graph.

```{r}
#| label: plot-aesthetic-mapping
ggplot(data = bechdel, mapping = aes(x = clean_test))
```

As we previously mentioned, we often omit the names of the first two arguments in R functions.
So you'll often see this written as:

```{r}
#| label: plot-simplified-call
ggplot(bechdel, aes(x = clean_test))
```

Note that the result is exactly the same.

-   The `geom_xx` function specifies the type of plot we want to use to represent the data. In the code below, we use `geom_point` which creates a plot where each observation is represented by a point.

```{r}
#| label: plot-point
ggplot(bechdel, aes(x = clean_test)) +
  geom_bar()
```

What types of movies are more common, those that pass or do not pass the test?

### Render, commit, and push

1.  Render your Quarto document.

2.  Go to the Git pane and check the box next to each file listed, i.e., stage your changes.
    Commit your staged changes using a simple and informative message.

3.  Click on push (the green arrow) to push your changes to your application exercise repo on GitHub.

4.  Go to your repo on GitHub and confirm that you can see the updated files.
    Once your updated files are in your repo on GitHub, you're good to go!

## Return-on-investment

Let's take a look at return-on-investment (ROI) for movies that do and do not pass the Bechdel test.

### Step 1 - Your turn

Create side-by-side box plots of `roi` by `clean_test` where the boxes are colored by `binary`.

```{r}
#| label: roi-zoom
# add code here
```

What are those movies with *very* high returns on investment?

```{r}
#| label: roi-outliers
bechdel |>
  filter(roi > 400) |>
  select(title, roi, budget_2013, gross_2013, year, clean_test)
```

### Step 2 - Demo

Expand on your plot from the previous step to zoom in on movies with `roi < ___` to get a better view of how the medians across the categories compare.

```{r}
#| label: roi-test-zoom
# add code here
```

What does this plot say about return-on-investment on movies that pass the Bechdel test?

### Render, commit, and push

1.  Render your Quarto document.

2.  Go to the Git pane and check the box next to each file listed, i.e., stage your changes.
    Commit your staged changes using a simple and informative message.

3.  Click on push (the green arrow) to push your changes to your application exercise repo on GitHub.

4.  Go to your repo on GitHub and confirm that you can see the updated files.
    Once your updated files are in your repo on GitHub, you're good to go!
