---
title: "AE 08: Durham climate + factors"
subtitle: "Suggested answers"
---

::: callout-important
These are suggested answers.
This document should be used as a reference only; it's not designed to be an exhaustive key.
:::

## Getting started

### Packages

We will use the **tidyverse** package in this application exercise.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
```

### Data

The data come from <https://www.usclimatedata.com/climate/durham/north-carolina/united-states/usnc0192> and are stored as `durham-climate.csv` in the `data` folder.

```{r}
#| label: load-data
#| message: false
durham_climate <- read_csv("data/durham-climate.csv")
```

And let's take a look at the data.

```{r}
#| label: view-data
durham_climate
```

## Reorder

What's wrong with the following plot?

Months are out of order.

```{r}
#| label: avg-high-broken
#| fig-width: 8
#| fig-asp: 0.4
#| out-width: 100%
ggplot(
  durham_climate, 
  aes(x = month, y = avg_high_f, group = 1)
  ) +
  geom_line() +
  geom_point(
    shape = "circle filled", size = 2,
    color = "black", fill = "white", stroke = 1
  ) +
  labs(
    x = "Month",
    y = "Average high temperature (F)",
    title = "Durham climate"
  )
```

Fix the plot.

```{r}
#| label: avg-high-fixed
#| fig-width: 8
#| fig-asp: 0.4
#| out-width: 100%
durham_climate <- durham_climate |>
  mutate(month = fct_relevel(month, month.name))

ggplot(
  durham_climate,
  aes(x = month, y = avg_high_f, group = 1)
  ) +
  geom_line() +
  geom_point(
    shape = "circle filled", size = 2,
    color = "black", fill = "white", stroke = 1
  ) +
  labs(
    x = "Month",
    y = "Average high temperature (F)",
    title = "Durham climate"
  )
```

## Recode and reorder

Update the plot above, coloring points based on season.
Additionally:

-   Make sure the legend is on top of the plot and the seasons appear in the legend in the order they appear in the plot (Winter, Spring, Summer, Fall).

-   Use `"circle filled"` as the `shape` for points, set their `size` to 3 points, outline (`stroke`) to 1 point, and fill them in with the following colors:

    -   Winter: `lightskyblue1`
    -   Spring: `chartreuse3`
    -   Summer: `gold2`
    -   Fall: `lightsalmon4`

```{r}
#| label: avg-high-seasons
#| fig-width: 8
#| fig-asp: 0.4
#| out-width: 100%
durham_climate <- durham_climate |>
  mutate(
    season = case_when(
      month %in% c("December", "January", "February") ~ "Winter",
      month %in% c("March", "April", "May") ~ "Spring",
      month %in% c("June", "July", "August") ~ "Summer",
      month %in% c("September", "October", "November") ~ "Fall",
    ),
    season = fct_relevel(season, "Winter", "Spring", "Summer", "Fall")
  )

ggplot(
  durham_climate,
  aes(x = month, y = avg_high_f, group = 1)
  ) +
  geom_line() +
  geom_point(
    aes(fill = season),
    shape = "circle filled", size = 3, stroke = 1
  ) +
  scale_fill_manual(
    values = c(
      "Winter" = "lightskyblue1",
      "Spring" = "chartreuse3",
      "Summer" = "gold2",
      "Fall" = "lightsalmon4"
    )
  ) +
  labs(
    x = "Month",
    y = "Average high temperature (F)",
    title = "Durham climate",
    fill = "Season"
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```

## Pivot

Overlay lines for both high and low temperatures on the same plot.
Additionally, use different colors for the two lines – a darker color for high and a lighter color for low temperatures.

```{r}
#| label: avg-seasons
#| fig-width: 8
#| fig-asp: 0.4
#| out-width: 100%
durham_climate |>
  pivot_longer(
    cols = c(avg_high_f, avg_low_f),
    names_to = "temp_type",
    names_prefix = "avg_",
    values_to = "avg_temp_f"
  ) |>
  mutate(temp_type = str_remove(temp_type, "_f")) |>
  ggplot(aes(x = month, y = avg_temp_f, group = temp_type, color = temp_type)) +
  geom_line() +
  geom_point(
    aes(fill = season),
    shape = "circle filled", size = 3, stroke = 1
  ) +
  scale_fill_manual(
    values = c(
      "Winter" = "lightskyblue1",
      "Spring" = "chartreuse3",
      "Summer" = "gold2",
      "Fall" = "lightsalmon4"
    )
  ) +
  scale_color_manual(
    values = c(
      "high" = "gray20",
      "low" = "gray70"
    )
  ) +
  labs(
    x = "Month",
    y = "Average temperature (F)",
    title = "Durham climate",
    fill = "Season",
    color = "Type"
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```
