---
title: "AE 14: Modelling penguins with multiple predictors"
subtitle: "Suggested answers"
categories: 
  - Application exercise
  - Answers
fig-width: 7
fig-asp: 0.618
---

::: callout-important
These are suggested answers.
This document should be used as reference only, it's not designed to be an exhaustive key.
:::

In this application exercise we will continue to study penguins.
The data can be found in the **palmerpenguins** package and we will use **tidyverse** and **tidymodels** for data exploration and modeling, respectively.

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
```

# Body mass vs. flipper length and island

Next, we will expand our understanding of models by continuing to learn about penguins.
So far, we modeled body mass by flipper length, and in a separate model, modeled body mass by island.
Could it be possible that the estimated body mass of a penguin changes by both their flipper length AND by the island they are on?

-   **Demo:** Fit a model to predict body mass from flipper length and island. Display the summary output and write out the estimated regression equation below.

```{r}
#| label: bm-fl-island-additive-fit
bm_fl_island_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + island, data = penguins)

tidy(bm_fl_island_fit)
```

$$
\widehat{body~mass} = -4625 + 44.5 \times flipper~length - 262 \times Dream - 185 \times Torgersen
$$

## Additive vs. interaction models

-   **Your turn:** Run the two chunks of code below and create two separate plots. How are the two plots different than each other? Which plot does the model we fit above represent?

```{r}
#| label: additive-interaction-viz
#| layout-ncol: 2
# Plot A
ggplot(
  penguins, 
  aes(x = flipper_length_mm, y = body_mass_g, color = island)
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  labs(title = "Plot A - Interaction model") +
  theme(legend.position = "bottom")

# Plot B
bm_fl_island_aug <- augment(bm_fl_island_fit, new_data = penguins)
ggplot(
  bm_fl_island_aug, 
  aes(x = flipper_length_mm, y = body_mass_g, color = island)
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(y = .pred), method = "lm") +
  labs(title = "Plot B - Additive model") +
  theme(legend.position = "bottom")
```

Plot B represents the model we fit.

-   **Your turn:** Interpret the slope coefficient for flipper length in the context of the data and the research question.

For every 1 millimeter the flipper is longer, we expect body mass to be higher, on average, by 44.5 grams, holding all else (the island) constant.
In other words, this is true for penguins in a given island, regardless of the island.

-   **Demo:** Predict the body mass of a Dream island penguin with a flipper length of 200 mm based on the additive model.

```{r}
#| label: dream-200-predict-additive
penguin_200_Dream <- tibble(
  flipper_length_mm = 200,
  island = "Dream"
)

predict(bm_fl_island_fit, new_data = penguin_200_Dream)
```

-   **Review:** Look back at Plot B. What assumption does the additive model make about the slopes between flipper length and body mass for each of the three islands?

The additive model assumes the same slope between body mass and flipper length for all three islands.

-   **Demo:** Now fit the interaction model represented in Plot A and write the estimated regression model.

```{r}
#| label: bm-fl-island-interaction-fit
bm_fl_island_int_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm * island, data = penguins)

tidy(bm_fl_island_int_fit)
```

$$
\widehat{body~mass} = -5464 \\
+ 48.5 \times flipper~length \\
+ 3551 \times Dream + 3218 \times Torgersen \\
- 19.4 \times flipper~length*Dream - 17.4 \times flipper~length*Torgersen
$$

-   **Review:** What does modeling body mass with an interaction effect get us that without doing so does not?

The interaction effect allows us to model the rate of change in estimated body mass as flipper length increases as different in the three islands.

-   **Your turn:** Predict the body mass of a Dream island penguin with a flipper length of 200 mm based on the interaction model.

```{r}
#| label: dream-200-predict-interaction
predict(bm_fl_island_int_fit, new_data = penguin_200_Dream)
```

## Choosing a model

Rule of thumb: **Occam's Razor** - Don't overcomplicate the situation!
We prefer the *simplest* best model.

```{r}
#| label: glance-models
glance(bm_fl_island_fit)
glance(bm_fl_island_int_fit)
```

-   **Review:** What is R-squared? What is adjusted R-squared?

R-squared is the percent variability in the response that is explained by our model.
(Can use when models have same number of variables for model selection)

Adjusted R-squared is similar, but has a penalty for the number of variables in the model.
(Should use for model selection when models have different numbers of variables).

# Your turn

-   Now, explore body mass, and it's relationship to bill length and flipper length.
    Brainstorm: How could we visualize this?

-   Fit the additive model.
    Interpret the slope for flipper in context of the data and the research question.

```{r}
#| label: bm-fl-bl-fit
bm_fl_bl_fit <- linear_reg() |>
  fit(body_mass_g ~ flipper_length_mm + bill_length_mm, data = penguins)

tidy(bm_fl_bl_fit)
```

Holding all other variables constant, for every additional millimeter in flipper length, we expect the body mass of penguins to be higher, on average, by 48.1 grams.
