---
title: "AE 10: Opinion articles in The Chronicle"
subtitle: "Suggested answers"
editor_options: 
  chunk_output_type: console
---

::: callout-important
These are suggested answers.
This document should be used as reference only, it's not designed to be an exhaustive key.
:::

## Part 1 - Data scraping

See [`chronicle-scrape.R`](https://github.com/sta199-f24/sta199-f24.github.io/blob/main/ae/_chronicle-scrape-A.R) for suggested scraping code.

## Part 2 - Data analysis

Let's start by loading the packages we will need:

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
```

-   Load the data you saved into the `data` folder and name it `chronicle`.

```{r}
#| label: load-data
#| message: false
chronicle <- read_csv("data/chronicle.csv")
```

-   Who are the most prolific authors of the 500 most recent opinion articles in The Chronicle?

```{r}
#| label: count-authors
chronicle |>
  count(author, sort = TRUE)
```

-   Draw a line plot of the number of opinion articles published per day in The Chronicle.

```{r}
#| label: articles-per-day
chronicle |>
  count(date) |>
  ggplot(aes(x = date, y = n, group = 1)) +
  geom_line()
```

-   What percent of the most recent 100 opinion articles in The Chronicle mention "climate" in their title?

```{r}
#| label: climate-title
chronicle |>
  mutate(
    title = str_to_lower(title),
    climate = if_else(str_detect(title, "climate"), "mentioned", "not mentioned")
    ) |>
  count(climate) |>
  mutate(prop = n / sum(n))
```

-   What percent of the most recent 100 opinion articles in The Chronicle mention "climate" in their title or abstract?

```{r}
#| label: climate-title-abstract
chronicle |>
  mutate(
    title = str_to_lower(title),
    abstract = str_to_lower(abstract),
    climate = if_else(
      str_detect(title, "climate") | str_detect(abstract, "climate"), 
      "mentioned", 
      "not mentioned"
      )
    ) |>
  count(climate) |>
  mutate(prop = n / sum(n))
```

-   Come up with another question and try to answer it using the data.

```{r}
#| label: add-label
# add code here
```
