---
title: Lab 2
subtitle: Revisiting the Midwest
categories: Lab
description: "Due Mon, Sep 23 at 8:30 am"
draft: false
---

# Introduction

In this lab, you'll continue to hone your data science workflow and integrate what you learned so far in the course (data visualization) with what's coming up (data wrangling).

::: callout-note
This lab assumes you've completed [Lab 0](/labs/lab-0.html) and [Lab 1](/labs/lab-1.html) and doesn't repeat setup and overview content from those labs.
If you haven't done those yet, you should review them before starting with this one.
:::

## Learning objectives

By the end of the lab, you will...

-   Be able to create transform data using `dplyr`
-   Build on your mastery of data visualizations using `ggplot2`
-   Get more experience with data science workflow using R, RStudio, Git, and GitHub
-   Further your reproducible authoring skills with Quarto
-   Improve your familiarity with version control using Git and GitHub

## Getting started

### Step 1: Log in to RStudio

-   Go to <https://cmgr.oit.duke.edu/containers> and log in with your Duke NetID and Password.
-   Click `STA198-199` under My reservations to log into your container. You should now see the RStudio environment.

### Step 2: Clone the repo & start a new RStudio project

-   Go to the course organization at [github.com/sta199-f24](https://github.com/sta199-f24) organization on GitHub.
    Click on the repo with the prefix **lab-2**.
    It contains the starter documents you need to complete the lab.

-   Click on the green **CODE** button and select **Use SSH** (this might already be selected by default; if it is, you'll see the text **Clone with SSH**).
    Click on the clipboard icon to copy the repo URL.

-   In RStudio, go to *File* ➛ *New Project* ➛*Version Control* ➛ *Git*.

-   Copy and paste the URL of your assignment repo into the dialog box *Repository URL*. Again, please make sure to have *SSH* highlighted under *Clone* when you copy the address.

-   Click *Create Project*, and the files from your GitHub repo will be displayed in the *Files* pane in RStudio.

-   Click *lab-2.qmd* to open the template Quarto file.
    This is where you will write up your code and narrative for the lab.

### Step 3: Update the YAML

In `lab-2.qmd`, update the `author` field to your name, render your document, and examine the changes.
Then, in the Git pane, click on **Diff** to view your changes, add a commit message (e.g., "Added author name"), and click **Commit**.
Then, push the changes to your GitHub repository and, in your browser, confirm that these changes have indeed propagated to your repository.

::: callout-important
If you encounter any issues with the above steps, flag a TA for help before proceeding.
:::

## Packages

In this lab, we will work with the **tidyverse** package, a collection of packages for performing data analysis in a "tidy" way.

```{r}
#| eval: true
#| message: false
library(tidyverse)
```

-   **Run** the code cell by clicking on the green triangle (play) button for the code cell labeled `load-packages`. This loads the package so that its features (the functions and datasets in it) are accessible from your *Console*.
-   Then, **render** the document that loads this package to make its features (the functions and datasets in it) available for other code cells in your Quarto document.

## Guidelines

As we've discussed in the lecture, your plots should include an informative title, axes and legends should have human-readable labels and aesthetic choices should be carefully considered.

Additionally, code should follow the [tidyverse style](https://style.tidyverse.org/).
Particularly,

-   there should be spaces before and line breaks after each `+` when building a `ggplot`,

-   there should also be spaces before and line breaks after each `|>` in a data transformation pipeline,

-   code should be properly indented,

-   there should be spaces around `=` signs and spaces after commas.

Furthermore, all code should be visible in the PDF output, i.e., should not run off the page on the PDF.
Long lines that run off the page should be split across multiple lines with line breaks.[^1]

[^1]: Remember, haikus, not novellas, when writing code!

As you complete the lab and other assignments in this course, remember to develop a sound workflow for reproducible data analysis.
This assignment will periodically remind you to render, commit, and push your changes to GitHub.

::: callout-important
You should have at least 3 commits with meaningful commit messages by the end of the assignment.
:::

# Questions

## Part 1

**All about `group_by()`!**

Use the following data frame for Question 1 and Question 2:

```{r}
#| eval: true
df <- tibble(
  var_1 = c(10, 20, 30, 40, 50),
  var_2 = c("Pizza", "Burger", "Pizza", "Pizza", "Burger"),
  var_3 = c("Apple", "Apple", "Pear", "Pear", "Banana")
)

df
```

### Question 1

Grouping by a single variable.

a\.
What does the following code chunk do?
Run it, analyze the result, and articulate in words what `arrange()` does.

```{r}
#| eval: false
df |>
  arrange(var_2)
```

b\.
What does the following code chunk do?
Run it, analyze the result, and articulate in words what `group_by()` does.
Also, comment on how it's different from the `arrange()` in part (a).

```{r}
#| eval: false
df |>
  group_by(var_2)
```

c\.
What does the following code chunk do?
Run it, analyze the result, and articulate in words what the pipeline does.

```{r}
#| eval: false
df |>
  group_by(var_2) |>
  summarize(mean_var_1 = mean(var_1))
```

### Question 2

Grouping by two variables.

a\.
What does the following code chunk do?
Run it, analyze the result, and articulate in words what the pipeline does.
Then, comment on what the message says.

```{r}
#| eval: false
df |>
  group_by(var_2, var_3) |>
  summarize(mean_var_1 = mean(var_1))
```

b\.
What does the following code chunk do?
Run it, analyze the result, and articulate in words what the pipeline does, especially mentioning what the `.groups` argument does.
How is the output different from the one in part (a)?

```{r}
#| eval: false
df |>
  group_by(var_2, var_3) |>
  summarize(mean_var_1 = mean(var_1), .groups = "drop")
```

c\.
What do the following pipelines do?
Run both, analyze their results, and articulate in words what each pipeline does.
How are the outputs of the two pipelines different?

```{r}
#| eval: false
df |>
  group_by(var_2, var_3) |>
  summarize(mean_var_1 = mean(var_1), .groups = "drop")

df |>
  group_by(var_2, var_3) |>
  mutate(mean_var_1 = mean(var_1))
```

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answers for Questions 1 and 2".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

## Part 2

**Back to the Midwest!**

In this part, you will revisit and build on some of your findings from Lab 1, where you explored the `midwest` data frame.
Remember that this data frame is bundled with the **ggplot2** package and is automatically loaded when you load the tidyverse package.
As a refresher, the data contains demographic characteristics of counties in the Midwest region of the United States.
You can read the documentation for the data set, including variable definitions, by typing `?midwest` in the Console or searching for `midwest` in the Help pane.

### Question 3

Calculate the number of counties in each state and display your results in descending order of number of counties.
Which state has the highest number of counties, and how many?
Which state has the lowest number, and how many?

::: callout-note
The number of counties in a state can change over time, so the values you see in this output may not be true today.
:::

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 3".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 4

While two counties in a given state can't have the same name, some county names might be shared across states.
A classmate says "Look at that, there is a county called \_\_\_ in each state in this dataset!" In a single pipeline, discover all counties that could fill in the blanks.
Your response should be a data frame with only the county names that could fill in the blank and the number of times they appear in the data.

::: callout-tip
You will want to use the `filter()` function in your answer, which requires a logical condition to describe what you want to filter for.
For example, `filter(x > 2)` means filter for values of `x` greater than 2, and `filter(y <= 3)` means filter for values of y less than or equal to 3.

The table below is a summary of logical operators and how to articulate them in R.

| operator      | definition                |
|---------------|---------------------------|
| `<`           | less than                 |
| `<=`          | less than or equal to     |
| `>`           | greater than              |
| `>=`          | greater than or equal to  |
| `==`          | exactly equal to          |
| `!=`          | not equal to              |
| `x & y`       | `x` AND `y`               |
| `x` \| `y`    | `x` OR `y`                |
| `is.na(x)`    | test if `x` is `NA`       |
| `!is.na(x)`   | test if `x` is not `NA`   |
| `x %in% y`    | test if `x` is in `y`     |
| `!(x %in% y)` | test if `x` is not in `y` |
| `!x`          | not `x`                   |
:::

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 4".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 5

Return to the following box plot of population densities where you were asked to identify at least one outlier.

```{r}
#| label: popdensity-boxplot
#| fig-width: 5
#| fig-height: 3
#| echo: false
#| eval: true

ggplot(midwest, aes(x = popdensity)) +
  geom_boxplot() +
  labs(
    title = "Population densities of midwest counties",
    x = "Population density"
  )
```

In this question, we want you to revisit this box plot and identify the counties described in each section.

a\.
The counties with a population density higher than 25,000.
Your code must use the `filter()` function.

b\.
The county with the highest population density.
Your code must use the `max()` function.

Answer using a single data wrangling pipeline for each part.
Your response should be a data frame with five columns: county name, state name, population density, total population, and area, in this order.
If your response has multiple rows, the data frame should be arranged in descending order of population density.

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 5".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 6

In Lab 1 you were also asked to describe the distribution of population densities.
The following is one acceptable description that touches on the shape, center, and spread of this distribution.
Calculate the values that should go into the blanks.

> The distribution of population density of counties is unimodal and extremely right-skewed.
> A typical Midwestern county has population density of \_\_\_\_ people per unit area.
> The middle 50% of the counties have population densities between \_\_\_ to \_\_\_ people per unit area.

::: callout-tip
Think about which measures of center and spread are appropriate for skewed distributions.
:::

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 6".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 7

Another visualization from Lab 1 was the following, which showed the proportion of urban counties in each state.

```{r}
#| label: metro-barplot
#| fig-width: 7
#| fig-height: 3
#| echo: false
#| eval: true
midwest <- midwest |>
  mutate(metro = if_else(inmetro == 1, "Yes", "No"))

ggplot(midwest, aes(x = state, fill = metro)) +
  geom_bar(position = "fill") +
  labs(
    title = "Do some states have more urban counties than others?",
    x = "State",
    y = "Proportion in metro area",
    fill = "In metro area?"
  )
```

Calculate these proportions in a single data pipeline.

::: callout-tip
Remember, you'll first need to create a new variable called `metro` which takes on the value `Yes` if the value of `inmetro` is 1, and `No` otherwise.
You can refer to [Lab 1](/labs/lab-1.html#question-6) if you need a refresher on how to do this.
:::

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 7".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 8

Return to the following scatter plot of percentage below poverty vs. percentage of people with a college degree, where the color and shape of points are determined by `state` where you were asked to identify at least one county that is a clear outlier by name.

```{r}
#| label: percollege-percbelowpoverty-scatterplot
#| fig-width: 5
#| fig-height: 3
#| echo: false
#| eval: true
ggplot(
  midwest,
  aes(
    x = percollege, y = percbelowpoverty,
    color = state, shape = state
  )
) +
  geom_point() +
  labs(
    title = "Percentage with a college degree versus poverty rate",
    subtitle = "by State in the Midwest",
    x = "% college educated",
    y = "% below poverty",
    color = "State", shape = "State"
  ) +
  annotate(
    geom = "point",
    x = 7.35, y = 48.5, color = "orange", shape = "square open", 
    size = 4.5, stroke = 2
  ) +
  annotate(
    geom = "point",
    x = 45, y = 10.2, color = "red", shape = "square open", 
    size = 25, stroke = 2
  ) +
  theme_minimal()
```

a\.
In a single pipeline, identify the observations marked in the orange square in the upper left corner.
Your answer should be a data frame with four variables: county, state, percentage below poverty, and percentage college educated.

b\.
In a single pipeline, identify the observations marked in the red square in the plot above.
Your answer should again be a data frame with four variables: county, state, percentage below poverty, and percentage college educated.

c\.
Bring your answers from part (a) and part (b) together!
In a single pipeline, and a single `filter()` statement, identify the observations marked in the red and orange square in the plot above.
Your answer should again be a data frame with four variables: county, state, percentage below poverty, and percentage college educated.

d\.
Create a new variable in `midwest` called `potential_outlier`.
This variable should take on the value:

-   `Yes` if the point is one the ones you identified in part (c), i.e., one of the points marked in the squares in the plot above.

-   `No` otherwise.

Then, display the updated `midwest` data frame, with county, state, percentage below poverty, percentage college educated, `potential_outlier` as the selected variables, arranged in descending order of `potential_outlier`.

e\.
Recreate the visualization above, i.e., a scatterplot of the percentage below poverty vs. the percentage of people with a college degree.
However, color the points by the newly created `potential_outlier` variable instead of the state.

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 8".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 9

a\.
In a single pipeline, calculate the total population for each state and save the resulting data frame as `state_population`.
Then, display the data frame, `state_population`, in descending order of total population.

b\.
Then, in a separate pipeline, calculate the proportion of the total population in each state.
Once again, display the results in descending order of proportion of population.

::: callout-tip
In answering parts (a) and (b), you'll create two new variables, one for the total population and the other for the proportion of total proportion.
Make sure to give them "reasonable" names – short but evocative.
:::

c\.
Which Midwestern state is most populous, and what percent of the Midwest population lives there?
Which is the least populous and what percent lives there?

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 9".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

### Question 10

Calculate the average percentage below poverty for each state and save the resulting data frame as `state_poverty` with the columns `state` and `mean_percbelowpoverty`.

Then, in a new pipeline, display the `state_poverty` data frame in ascending order of `mean_percbelowpoverty`.
Which state has the lowest average percentage below poverty across its counties?
Which state has the highest average percentage below poverty across its counties?

::: render-commit-push
Render, commit, and push your changes to GitHub with the commit message "Added answer for Question 10".

Make sure to commit and push all changed files so that your Git pane is empty afterward.
:::

# Wrap-up

## Submission

Once you are finished with the lab, you will submit your final PDF document to Gradescope.

::: callout-warning
Before you wrap up the assignment, make sure all of your documents are updated on your GitHub repo.
We will be checking these to make sure you have been practicing how to commit and push changes.

To be considered " on time, " you must turn in a PDF file to the Gradescope page by the submission deadline.
:::

To submit your assignment:

-   Go to <http://www.gradescope.com> and click *Log in* in the top right corner.
-   Click *School Credentials* $\rightarrow$ *Duke NetID* and log in using your NetID credentials.
-   Click on your *STA 199* course.
-   Click on the assignment, and you'll be prompted to submit it.
-   Mark all the pages associated with each question. All the pages of your lab should be associated with at least one question (i.e., should be "checked").

::: callout-important
## Checklist

Make sure you have:

-   attempted all questions
-   rendered your Quarto document
-   committed and pushed everything to your GitHub repository such that the Git pane in RStudio is empty
-   uploaded your PDF to Gradescope
-   selected pages associated with each question on Gradescope
:::

## Grading and feedback

-   Some of the questions will be graded for accuracy.

-   Some will be graded for completion.

-   There are also workflow points, for coding style, for committing at least three times as you work through your lab, and for overall organization.

-   You'll receive feedback on your lab on Gradescope within a week.
